Towards a Universal Facebook of the Ancient World

### 2016-01-20 source/18/EAGLE2016_Full_Paper_Broux.tex
further changes
Bronx not moved to pdfs folder, as it is being compiled by Francesco

To reconstruct proper social networks, a decent prosopography is indispensable. In a traditional scholarly setting, this
implies time-consuming and painstaking manual labor. Fortunately, in a digital environment, there are other ways, and
Trismegistos (www.trismegistos.org) forms an ideal starting point. Trismegistos grew out of a long tradition of
databases and prosopographies (structured lists of people), as well as various other Ancient History projects at KU
Leuven. The original idea of Trismegistos was to foster interdisciplinarity in the study of Ancient Egyptian society by
creating a central database with metadata about published papyrological texts from Greco-Roman Egypt, in a first
instance written in Greek, Latin and Egyptian (including hieroglyphic, hieratic and demotic). The inclusion of Egyptian
soon dissolved the disciplinary boundary with epigraphy, broadened the chronological window, which was eventually set
to 800 BC – AD 800, and led to the inclusion of further languages such as Coptic, Aramaic and Arabic.

@@

Apart from this gazetteer, rules were developed to cope with the combination of names, or more correctly declined name
variants, as well as with the combination of names with non-onomastic introductory terms (e.g. `son of') in the
identification of individuals. For Greek texts, this is pretty straightforward, as people were identified by their name
(generally a single name, in some cases a double name) followed by genealogical identifiers (in the genitive form)
only. The onomastic habits encountered in Latin inscriptions differ significantly, however. Here the majority of the
individuals follow the Roman naming system. This consists of a fossilized praenomen, a nomen gentilicium (the
equivalent of our family name), and a personal cognomen. Often the patronymic was inserted between the gentilicium and
cognomen; sometimes even the papponymic and the voting tribe, e.g. Marcus Tullius Marci filius Marci nepos Cornelia
Cicero (`Marcus Tullius Cicero, son of Marcus, grandson of Marcus, [of the] Cornelia [tribe]'). A completely new set of
rules is therefore being developed to apply to the Latin inscriptions.

>>>>

To reconstruct proper social networks, a decent prosopography is indispensable. In a traditional scholarly setting, this
implies time-consuming and painstaking manual labor. Fortunately, in a digital environment, there are other ways, and
Trismegistos (\url{www.trismegistos.org}) forms an ideal starting point. Trismegistos grew out of a long tradition of
databases and prosopographies (structured lists of people), as well as various other Ancient History projects at KU
Leuven. The original idea of Trismegistos was to foster interdisciplinarity in the study of Ancient Egyptian society by
creating a central database with metadata about published papyrological texts from Greco-Roman Egypt, in a first
instance written in Greek, Latin and Egyptian (including hieroglyphic, hieratic and demotic). The inclusion of Egyptian
soon dissolved the disciplinary boundary with epigraphy, broadened the chronological window, which was eventually set
to 800 BC – AD 800, and led to the inclusion of further languages such as Coptic, Aramaic and Arabic.

@@

Apart from this gazetteer, rules were developed to cope with the combination of names, or more correctly declined name
variants, as well as with the combination of names with non-onomastic introductory terms (e.g. `son of') in the
identification of individuals. For Greek texts, this is pretty straightforward, as people were identified by their name
(generally a single name, in some cases a double name) followed by genealogical identifiers (in the genitive form)
only. The onomastic habits encountered in Latin inscriptions differ significantly, however. Here the majority of the
individuals follow the Roman naming system. This consists of a fossilized praenomen, a nomen gentilicium (the
equivalent of our family name), and a personal cognomen. Often the patronymic was inserted between the gentilicium and
cognomen; sometimes even the papponymic and the voting tribe, e.g. \emph{Marcus Tullius Marci filius Marci nepos Cornelia
Cicero} (`Marcus Tullius Cicero, son of Marcus, grandson of Marcus, [of the] Cornelia [tribe]'). A completely new set of
rules is therefore being developed to apply to the Latin inscriptions.
